Title: Data Transformation with R 
Author : Jedlueng 
Source: Data Transformation - Data Rockie 

#Install Library 
```{r}
library(tidyverse)
library(jsonlite)
library(RSQLite)
library(RPostgreSQL)
library(sqldf)
```



Data Transformation is essential for machine learning and data analytics 


- dplyr -> grammar of data manipulation 
- using mtcars 


#Viewing data 

##Show rows name 

```{r}
view(mtcars) #View data 
rownames(mtcars) #show all rows 

```

##Show column name 
```{r}
colnames(mtcars)
```

#DPLYR 
5 functions 
1. select() 
2. filter() 
3. arrange() 
4. mutate()
5. summarise() + group_by() 


Comparing with SQL 
5 functions 
1. select() => SELECT
2. filter() => WHERE
3. arrange() => ORDER BY 
4. mutate() => SELECT .. AS newColumn 
5. summarise() + group_by() 

##SELECT 
```{r}
select(mtcars, 1:5) #select column 1-5
```

```{r}
select(mtcars, 1:3, 5, 10)
```


### SELECT Using column name 


```{r}
select(mtcars, mpg, wt, hp)
```




###SELECT both name and index 
```{r}
select(mtcars, mpg, wt, hp , 10:11)
```


###SELECT started with A 
```{r}
select(mtcars, starts_with("a"))
```

###SELECT ends_with()
```{r}
select(mtcars, ends_with("p"))
```

###SELECT contains() 
```{r}
select(mtcars, contains("a"))
```



###select everything() 
```{r}
select(mtcars, carb, everything())
```

#Making new column 
```{r}
mtcars$model <- rownames(mtcars) #model is a new column 
#Remove rowname 
rownames(mtcars) <- NULL

select(mtcars, model, everything())
```




#Data Pipeline %>% (Pipe operator)

##Bring data out of the function 
```{r}
select(mtcars, 1:5)

mtcars %>% select(1:5)

#Both of these are the same 
```
```{r}
mtcars %>% 
  select(mpg,hp,wt) %>% 
  filter(hp < 100)
```


##Filter() 
```{r}
mtcars %>% 
  select(mpg,hp,wt) %>% 
  filter(hp <100, wt < 2)

mtcars %>% 
  select(mpg,hp,wt) %>% 
  filter(hp <100| wt < 2)
```
##Arrage()

##มากไปน้อย
```{r}
mtcars %>% 
  select(mpg,hp,wt) %>% 
  filter(hp<100|wt <2)%>%
  arrange(hp)
```
##น้อยไปมาก

```{r}
mtcars %>% 
  select(mpg,hp,wt) %>% 
  filter(hp<100|wt <2)%>%
  arrange(desc(hp))
```



##Filter data == WHERE SQL 

```{r}
mtcars %>% 
  select(model, mpg, hp, wt,am) %>% 
  filter(mpg > 25, mpg <=30 ) #More than 25 less than or equal to 30 


mtcars %>% 
  select(model, mpg, hp, wt,am) %>% 
  filter(mpg %in% 25:30) #More than 25 less than or equal to 30 
```


```{r}
mtcars %>% 
  select(model, mpg,hp,wt,am) %>% 
  filter(grepl("^Mazda", model))
```


#Mutate() 
```{r}
mtcars %>% 
  select(model,mpg,hp,wt,am)%>%
  mutate(hello = "hello") #Create new column hello 
```

##Using if statement 
```{r}
mtcars %>% 
  select(model,mpg,hp,wt,am) %>%
           mutate(hp_segment = if_else(hp<100,"low","high")) %>%
  head(10)
```


##CASE WHEN 
Create a new column 

```{r}
m2 <- mtcars %>% 
  select(model,mpg,hp,wt,am) %>% 
  mutate(hp_segment = if_else(hp<100,"low","high"),
         hp_segment2 = case_when(
           hp<100 ~"low",
           hp <200 ~ "medium",
           TRUE ~"high"
         ))
print(head(m2))
```

#Show data type 
```{r}
str(mtcars)
```


```{r}
glimpse(mtcars)
```


#Change Binary to String 

```{r}
am <- mtcars %>% 
  select(model,am) %>%
  mutate(am = if_else(am == 0, "Auto", "Manual")) %>%
  head(10) 
```

##Mutiple changed 
```{r}
mtcars <- mtcars %>%
  mutate(am = if_else(am == 0, "Auto", "Manual"), 
         vs = if_else(vs == 0, "V-Shaped", "Straight"))
```


#count 
```{r}
mtcars %>% 
  count(am)
```

##Count multiple column 
```{r}
mtcars %>% 
  count(am,vs) 
```




```{r}
m3 <- mtcars %>% 
  count(am,vs ) %>% 
  mutate(percent = n/nrow(mtcars))
```

#Read Write CSV Files 
```{r}
write_csv(m3, "summary_mtcars.csv")
#csv file will be in the current working directory
```



##Read csv 
```{r}
m3 <- read_csv("summary_mtcars.csv")
m3
```






#Data Type 


##Change string to factor 
```{r}
mtcars <- mtcars %>% 
  select(model,mpg,vs,am) %>% 
  mutate(vs = as.factor(vs),
         am = as.factor(am))

print(mtcars)
```


#SSummary 

```{r}
mtcars %>% 
  group_by(am, vs) %>%
  summarise(
    avg_mpg = mean(mpg),
    sum_mpg = sum(mpg),
    min_mpg = min(mpg),
    max_mpg = max(mpg),
    var_mpg = var(mpg), 
    n = n()
  )
```


#Join table 

both of these table can be join together 
(PAUL)
```{r}
band_members
```


```{r}
band_instruments
```

##inner join 

```{r}
inner_join(band_members,
           band_instruments,
           by = "name")


```

```{r}
left_join(band_members,
          band_instruments,
          by = "name")
```

##Right Join 
```{r}
right_join(band_members, 
           band_instruments,
           by = "name")
```



##Full Join 

```{r}
full_join(band_members, 
          band_instruments,
          by = "name")
```


##Refactor 

```{r}
band_members %>% 
  full_join(band_instruments, 
            by = "name") %>% 
  filter(name %in% c("Join", "Paul")) %>% 
  mutate(hello = "OK")
```



Maximum 8-10 line > else too much 

#NYCflights 13 

```{r}
library(nycflights13)
```

```{r}
flights
```

```{r}
view(flights)
```

```{r}
flights %>%
  filter(month == 9) %>%
  count(origin,dest)
```

Questions 
1. carriers most flight 
2. in March - May 2013 
3. Origin == JFK 



```{r}
df <-flights %>% 
  filter(origin == "JFK" & month %in% c(3,4,5)) %>%
  count(carrier) %>% 
  arrange(desc(n))
```

##Sent back to users. 
```{r}
write_csv(df,"requested_data.csv")
```


#Mock up data Create new data 

```{r}
students <- data.frame( 
  id = 1:5, 
  name = c("toy", "joe", "anna", "mary", "kevin"),
  cid = c(1,2,2,3,2))
```



```{r}
course <- data.frame(
  course_id = 1:3 ,
  course_name = c("Data", "R", "Python"
))
```

##Left Join 
```{r}
students %>% 
  left_join(course, by = c("cid" = "course_id" )) %>%  filter(course_name == "R")
```


30 functions is essential for data analyst in R 

#University (JOIN many tables)
```{r}
university <- data.frame(
  uid = 1:2,
  uname = c("University of London", "Chula University")
)
```

##Join more than two tables
```{r}
students %>% 
  left_join(course,by = c("cid"= "course_id")) %>% left_join(university, by = "uid")
```



#Wide -> Long Format Transformation 

```{r}
long_worldphones <- WorldPhones %>% 
  as.data.frame() %>% 
  rownames_to_column(var = "year")%>% 
  pivot_longer(N.Amer:Mid.Amer,
               names_to = "Region",
               values_to = "Sales")

print(long_worldphones)
```

```{r}
long_worldphones %>% 
  group_by(Region) %>%
  summarise(total_sales = sum(Sales))
```


##Long -> Wide format 

```{r}
wide_data <- long_worldphones %>% 
  pivot_wider(names_from = "Region",
              values_from = "Sales") 
```


##Write 
```{r}
write_csv(wide_data, "data.csv")
```


#Chinook.db 
Connect SQL database 
```{r}
library(RSQLite)
library(RPostgreSQL)
```


##Connect database 
##Create connection > query > close con 
```{r}
#Connnect sql server 
con <- dbConnect(SQLite(),"chinook.db")
```

```{r}
#Show table 
dbListTables(con)
```

```{r}
#Select 
dbGetQuery(con, "SELECT 
                  firstname,
                  lastname,
                  country 
                FROM customers 
                WHERE country IN ('France', 'Austria', 'Belgium')")

```

##Join using SQLs 
```{r}
query01 <- dbGetQuery(con, "SELECT * FROM artists
                      JOIN albums ON artists.artistid = albums.artistid
                      JOIN tracks ON tracks.albumid = albums.albumid")
```


```{r}
tracks <- query01
```

```{r}
tracks %>% 
  select(Composer, Milliseconds, Bytes, Unitprice) %>% 
  head()
```

##Disconenct SQLs 
```{r}
dbDisconnect(con)
con
```


#Sample data (Janitor)

```{r}
library(janitor)
tracks_clean <- clean_names(tracks)
#Clean everything (Duplicate and whatever)
```


```{r}
set.seed(1)
tracks_clean %>% 
  select(1:4) %>% 
  sample_n(10) %>%
  head(10)
```

